Estimating duration in the presence of misclassification

### 2024-01-23 transmission.tex
backcalc plots

A slow increase in incidence occurs over the autumn 2020 period, peaking in late November in most regions (see \cref{transmission:fig:backcalc-regions}).
An exception is the North East and East of England regions (see \cref{transmission:fig:backcalc-regions}(C) and (D) respectively); these regions show no consistent increase or decrease in incidence during October.
All regions then decrease during November, coinciding with the national lockdown.
A rise then occurs, although starting at different times in different regions.
This could be due to the emergence of the Alpha variant, which was more transmissible than previous variants and emerged at different times across the country\todo{cite Alpha variant}.
This leads to a second peak in late December or early January in all regions; this second peak is larger than the first in all regions except the Northern three (North East, North West, and Yorkshire and the Humber).
\begin{figure}
    \todo[inline]{Four-panel regions figure}
    \capton[Incidence estimated using backcalculation by region]{%
        Incidence proportion for all of England (A) and its component regions.
        Regions are split into broad groupings of three regions each: Northern (B), Midlands (C), and Southern (D).
        Yorkshire and the Humber is labelled as Yorkshire for brevity.
    }
    \label{transmission:fig:backcalc-regions}
\end{figure}

The incidence proportion was highest among the 11--15 and 16--25 age groups (see \cref{transmission:fig:backcalc-ages}).
Except the 16--25 age group, all age groups show similar trends (\ie parallel lines on a log scale).
This is the pattern predicted by an age-stratified transmission model (see \cref{E-SEIR:sec:structured-populations}), if behavioural changes are independent of age group, and there is negligible impact of immunity.
The 16--25 age group decreases faster during November.
There are many plausible explanations for this effect.
The higher incidence in this age group prior to November means that this could be an immunity effect.
Alternatively, it could be a consequence of schools remaining open in the November lockdown, which has a small effect on 16--25 as they interact less with school-age children than other age groups.
Many other explanations are possible, and further work is required to determine causality.
\begin{figure}
    \todo[inline]{age groups figure}
    \capton[Incidence estimated using backcalculation by age group]{%
        Incidence proportion for each age group, on both natural and log scales.
    }
    \label{transmission:fig:backcalc-ages}
\end{figure}

An edge effect where a rapid rise is seen between day 1 and 2 is visible (see \cref{transmission:fig:start-effect}).
This is due to the incorrect assumption that incidence is constant in the past.
Changing this assumption (\eg to exponential growth) modifies the estimated value on the first day, but the rest of the series is insensitive to this change (not shown).
Therefore, the first day's estimate is unreliable, and has been excluded from all other plots.
However, the rest of the series is robust to this assumption and hence reliable.
\begin{figure}
    \todo[inline]{start effect figure}
    \capton[Edge effects in backcalculation method]{%
        First seven days of incidence proportion estimate for the whole of England, highlighting the edge effect.
    }
    \label{transmission:fig:start-effect}
\end{figure}

>>>>

A slow increase in incidence occurs over the autumn 2020 period, peaking in late November in most regions (see \cref{transmission:fig:backcalc-regions}).
An exception is the North East and East of England regions (see \cref{transmission:fig:backcalc-regions}(C) and (D) respectively); these regions show no consistent increase or decrease in incidence during October.
All regions then decrease during November, coinciding with the national lockdown.
A rise then occurs, although starting at different times in different regions.
This could be due to the emergence of the Alpha variant, which was more transmissible than previous variants and emerged at different times across the country\todo{cite Alpha variant}.
This leads to a second peak in late December or early January in all regions; this second peak is larger than the first in all regions except the Northern three (North East, North West, and Yorkshire and the Humber).
\begin{figure}
    \includegraphics{transmission/backcalc-regions}
    \caption[Incidence estimated using backcalculation by region]{%
        Incidence proportion for all of England (A) and its component regions.
        Regions are split into broad groupings of three regions each: Northern (B), Midlands (C), and Southern (D).
        Yorkshire and the Humber is labelled as Yorkshire for brevity.
    }
    \label{transmission:fig:backcalc-regions}
\end{figure}

The incidence proportion was highest among the 11--15 and 16--25 age groups (see \cref{transmission:fig:backcalc-ages}).
Except the 16--25 age group, all age groups show similar trends (\ie parallel lines on a log scale).
This is the pattern predicted by an age-stratified transmission model (see \cref{E-SEIR:sec:structured-populations}), if behavioural changes are independent of age group, and there is negligible impact of immunity.
The 16--25 age group decreases faster during November.
There are many plausible explanations for this effect.
The higher incidence in this age group prior to November means that this could be an immunity effect.
Alternatively, it could be a consequence of schools remaining open in the November lockdown, which has a small effect on 16--25 as they interact less with school-age children than other age groups.
Many other explanations are possible, and further work is required to determine causality.
\begin{figure}
    \makebox[\textwidth][c]{\includegraphics[width=.9\paperwidth]{transmission/backcalc-ages}}
    \caption[Incidence estimated using backcalculation by age group]{%
        Incidence proportion for each age group, on both natural and log scales.
    }
    \label{transmission:fig:backcalc-ages}
\end{figure}

An edge effect where a rapid rise is seen between day 1 and 2 is visible (see \cref{transmission:fig:start-effect}).
This is due to the incorrect assumption that incidence is constant in the past.
Changing this assumption (\eg to exponential growth) modifies the estimated value on the first day, but the rest of the series is insensitive to this change (not shown).
Therefore, the first day's estimate is unreliable, and has been excluded from all other plots.
However, the rest of the series is robust to this assumption and hence reliable.
\begin{figure}
    \centering \includegraphics{transmission/backcalc-start-effect}
    \caption[Edge effects in backcalculation method]{%
        First seven days of incidence proportion estimate for the whole of England, highlighting the edge effect.
    }
    \label{transmission:fig:backcalc-start-effect}
\end{figure}
